### 2016-02-16 paper/paper.tex
Add first version of paper

\documentclass[onecolumn, a4paper, 10pt]{article}

%%% Packages
\usepackage{graphicx}
\usepackage{balance}
\usepackage{tikz}
\usepackage{standalone}
\usepackage{colortbl}
\usepackage{subcaption}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{multirow,bigdelim}
\usepackage[pdftex, letterpaper, lmargin = 1.15in, rmargin = 1.15in]{geometry}

%%% TikZ libraries
\usetikzlibrary{backgrounds, positioning, decorations.pathreplacing, calc, fit}
\usetikzlibrary{shapes}

%%% Macros
\newcommand{\tbr}{{\color{red}\textbf{[TO BE REVISED]}}}
\newcommand{\tbd}{{\color{red}\textbf{[TO BE DONE]}}}

%\setlength{\parskip}{1em}

\begin{document}

% Title
\title{
  Spanner: Google's Globally-Distributed Database \newline
  Seminar Report
}

\author{
  Daniel Kocher \\
  Department of Computer Sciences \\
  University of Salzburg \\                                                    
  Daniel.Kocher@stud.sbg.ac.at
}

\maketitle

\section{Introduction \tbr}
\label{sec:introduction}

In the original paper~\cite{Corbett:2012}, Google introduced a
globally-distributed database referred to as \emph{Spanner}. This system is based
upon Paxos~\cite{Lamport:1998} state machines on machines spread across multiple
datacenters across the world. This global replication guarantees high availability
and also implies some desirable locality properties which benefit the performance.

Google already has two other highly scalable distributed databases, i.e.
Bigtable~\cite{Chang:2008} and Megastore~\cite{Baker:2011}. Hence, there must be
a reason, why they did not just use one of these but implemented another
distributed database. According to Google, it is difficult for applications/users
to deal with complex and evolving schemas. Moreover, Bigtable only guarantees to
be eventually-consistent. In contrast, Spanner provides much stronger consistency
guarantees. On the other hand, Megastore suffers from poor write throughput even
though it is used for many applications.

Therefore, Google decided to introduce another distributed, highly available
database system which evolved from a key-value store somehow similar to the one
used by Bigtable (but we will see that this is only partially correct). A core
feature provided by Spanner are so-called \emph{externally-consistent} distributed
general-purpose transactions. Of course, Spanner also provides an SQL-like query
language to enable applications/users to interact with Spanner in a well-known
way. The data is stored in schematized \emph{semi-relational} tables and the data
is versioned automatically. Google also implemented a garbage collection to deal
with old data. The replication can be dynamically configured in order to control
the locality of the replicas. This is useful to control write and read latency
as well as availability. Another main contribution of Spanner are
\emph{globally-consistent} reads at a specific timestamp. This and
externally-consistent reads and writes enable Spanner to provide a special type
of transaction, so-called \emph{schema-change} transactions, which provides atomic
schema changes without locking the whole database and all ongoing transactions.

To provide all these features, Google designed a novel \emph{TrueTime} API which
exposes clock uncertainty and enables Spanner to assign commit timestamps to
transactions which reflect serialization order.

\section{Implementation \tbr}
\label{sec:implementation}

A Spanner deployment is referred to as a \emph{universe} which consists of
multiple zones. A universe has two global components: A \emph{universemaster}
which, essentially, is a debugging console and shows the status of each zone. A
\emph{placement driver} which automatically moves data between zones due to
modified replication constraints or load balancing between zones.

Each zone is controlled by a \emph{zonemaster}. A zonemaster is responsible to
organize the data placement in a zone. The data is stored on so-called
\emph{spanservers} and the zonemaster assigns data to these spanservers. A zone
consists of multiple such spanservers which serve the data to clients. Spanner
somehow needs to enable the client to locate the spanservers which serve the data.
For this purpose, each zone contains multiple \emph{location proxies}.

According to Google, a zone can be seen as a Bigtable server deployment. The set
of zones also determines the locations across which the data can be replicated.
One can also think of a zone as geographically wide-spread locations, e.g.
continents. Using such a physical isolation, Spanner is resistent to events which
affect large areas, e.g. natural disasters or war. Hence, a well-configured
Spanner deployment is available almost all the time.
Figure~\ref{subfig:spanner-universe} illustrates the organization of a Spanner
universe.

\begin{figure}[ht]
  \centering
  \begin{subfigure}{0.45\textwidth}
    \includegraphics[width = \textwidth]{figs/spanner-server-organization.png}
    \caption{Spanner universe.~\cite{Corbett:2012}}
    \label{subfig:spanner-universe}
  \end{subfigure}
  \hfill
  \begin{subfigure}{0.45\textwidth}
    \includegraphics[width = \textwidth]{figs/spanserver-software-stack.png}
    \caption{Spanserver software stack.~\cite{Corbett:2012}}
    \label{subfig:spanserver-software-stack}
  \end{subfigure}
  \caption{Spanner hierarchy.}
  \label{fig:spanner-hierarchy}
\end{figure}

\subsection{Spanserver Software Stack \tbr}
\label{subsec:spanserver-software-stack}

Each spanserver also consists of multiple software layers (four at least). At the
bottom layer, every spanserver is responsible for between 100 and 1,000
\emph{tablets}. A tablet is bag of key-value mappings of the form
\begin{center}
  \texttt{(key:string, timestamp:int64)} $\rightarrow$ \texttt{value}
\end{center}
This is somehow similar to the tablet abstraction used in Bigtable. However,
Spanner \emph{automatically} assigns timestamps to data which makes Spanner a
\emph{multi-version} database. The state of a tablet is stored on a distributed
file system called \emph{Colossus}, the successor of the Google File
System~\cite{Ghemawat:2003}.

On top of each tablet, a Paxos state machine is implemented in order to support
consistent replication. One Paxos state machine is elect as leader. Leases are
assigned in a long-lived manner (default lease length is 10 seconds). A leader
implements three additional layers. A \emph{lock table} which stores the state
for two-phase locking and therefore implements the concurrency control. On top
of the lock table, a \emph{transaction manager} is implemented to coordinate
distributed transactions. In turn, this transaction manager is used to implement
a \emph{participant leader}. This leader is only utilized if a transaction
involves multiple Paxos groups (set of replicas). In this case, one of the
participant groups is chosen as coordinator and the leader of this group is
referred to as \emph{coordinator leader}.
Figure~\ref{subfig:spanserver-software-stack} shows the spanserver software stack
as described before.

\subsection{Directories \tbr}
\label{subsec:directories}

The next abstraction in the hierarchy is a bucketing abstraction. Google calls
this abstraction \emph{directory}. A directory is a set of contiguous keys.
All keys in a directory share a common prefix.

A directory is the unit of data placement/movement, hence only whole directories
are moved between Paxos groups, even across zones. Figure~\ref{subfig:directories}
illustrates this principle. Directory 2 is moved from Paxos group 1 to Paxos group
2 and at the same time from zone 2 to zone 3. Google shortly mentions that this
is an over-simplified view because a directory is split into multiple
\emph{fragments} if it grows too large. However, Google does not provide more
details.

All the data within a directory has the same replication configuration. As will
be described later, the concept of directories enables Google to improve the
performance of Spanner by placing directories in the same Paxos group which are
frequently accessed together. The fact that in general multiple directories are
placed in the same Paxos group is another difference between the key-value mapping
of Spanner and the one used in Bigtable. A Spanner tablet may contain multiple
contiguous ranges of keys (= directories) as opposed to a Bigtable tablet which
may only contain one contiguous range of keys. Figure~\ref{fig:comparison-tablets}
shows an example comparison of the two tablet types.

\begin{figure}[ht]
  \centering
  \begin{subfigure}{0.45\textwidth}
    \centering
    \begin{tabular}{|l||ll}
      \hline
      {\bfseries Key} & {\bfseries Value} \tabularnewline
      \hline\hline
      12341 & \ldots & \rdelim\}{3}{5mm}[Directory 1234] \tabularnewline
      12342 & \ldots \tabularnewline
      12343 & \ldots \tabularnewline
      \hline
      67891 & \ldots & \rdelim\}{4}{3mm}[Directory 6789] \tabularnewline
      67892 & \ldots \tabularnewline
      67893 & \ldots \tabularnewline
      67894 & \ldots \tabularnewline
      \hline
    \end{tabular}
    \caption{Spanner tablet}
    \label{subfig:spanner-tablet}
  \end{subfigure}
  \qquad
  \begin{subfigure}{0.45\textwidth}
    \centering
    \begin{tabular}{|l||l}
      \hline
      {\bfseries Key} & {\bfseries Value} \tabularnewline
      \hline\hline
      12341 & \ldots \tabularnewline
      12342 & \ldots \tabularnewline
      12343 & \ldots \tabularnewline
      12344 & \ldots \tabularnewline
      12345 & \ldots \tabularnewline
      12346 & \ldots \tabularnewline
      12347 & \ldots \tabularnewline
      \hline
    \end{tabular}
    \caption{Bigtable tablet}
    \label{subfig:bigtable-tablet}
  \end{subfigure}
  \caption{Comparison of Spanner's and Bigtable's tablets}
  \label{fig:comparison-tablets}
\end{figure}

This partitioning of the row space is critical for the performance of Spanner
because it allows Google to physically \emph{co-locate} data that is frequently
accessed together. Therefore, directories are the smallest unit for which an
application/user can specify geographic-replication properties. An application
can control replication in two dimension, i.e. the number and types of replicas,
and, the geographic placement of replicas.

\begin{figure}[ht]
  \centering
  \begin{subfigure}{0.45\textwidth}
    \includegraphics[width = \textwidth]{figs/directories}
    \caption{Directories (unit of data movement).~\cite{Corbett:2012}}
    \label{subfig:directories}
  \end{subfigure}
  \qquad
  \begin{subfigure}{0.45\textwidth}
    \includegraphics[width = \textwidth]{figs/interleaving-example.png}
    \caption{Two interleaved tables.~\cite{Corbett:2012}}
    \label{subfig:interleaved-tables}
  \end{subfigure}
  \caption{Directories and interleaved tables.}
  \label{fig:directories-and-tables}
\end{figure}

\subsection{Data Model \tbr}
\label{subsec:data-model}

There exists another layer on top of the directory-bucketed key-value mappings.
This layer is probably the most intuitive one for people who are used to common
database systems. The data model allows applications to create databases in a
universe each of which may contain any number of schematized
\emph{semi-relational} tables. A table has rows, columns and, as a third
dimension, versioned values. Spanner's tables are semi-relational because rows
must have names, i.e. every table must have an ordered set of one or more
primary-key columns. This is useful for applications because the locality of
data can be controlled by choosing keys accurately.

Therefore, Spanner requires the application/user to partition databases into
one or more hierarchies of tables. Based on this hierarchical information, Spanner
builds the directories. A directory contains hierarchically ordered rows from
all tables of a hierarchy which belong to the same key of the topmost table of
the hierarchy. Figure~\ref{subfig:interleaved-tables} shows an examples of this
concept. In this example, the \texttt{Users} table is the topmost table of the
hierarchy and the \texttt{Albums} table is one level below. As we can see, all
rows of \texttt{Albums} starting with \texttt{1} are located in the same directory
(\texttt{3365}). This directory has a common prefix, i.e. \texttt{1}, from the
\texttt{Users} table. The same principle can be seen for key \texttt{2} of the
\texttt{Users} table. This interleaving of tables exploits locality relationships
of multiple tables and therefore is significant for Spanner's performance.

\section{TrueTime \tbr}
\label{sec:truetime}

This section describes the key enabler of Spanner, i.e. the novel \emph{TrueTime}
API. Essentially, this API enables Google to provide timestamps for transactions
which reflect serialization order.

TrueTime represents time as a $TTinterval$. A $TTinterval$ is an interval with a
bounded time uncertainty and has two endpoints, i.e. $earliest$ and $latest$,
which are of type $TTstamp$ (which is a timestamp in the context of Spanner).
Table~\ref{tbl:truetime-api} summarizes the three methods provided by the TrueTime
API. However, the most important method is $TT.now()$ which returns a $TTinterval$
which is guaranteed to contain the absolute time of the invocation of $TT.now()$,
i.e. $TT.now().earliest \leq t_{abs}(e) \leq TT.now().latest$, where $e$ denotes
the event of invoking $TT.now()$. More precisely, the absolute time of invocation
of $TT.now()$ is exactly in the middle this $TTinterval$ and spans over
$2\cdot\epsilon$ where $\epsilon$ denotes the instantaneous error bound. $\epsilon$
is derived from applied worst-case local clock drift (assumed to be
$200\mu$s/sec). This instantaneous error bound is adjusted over time for every
single machine.

TrueTime uses GPS and atomic clocks as time references. Google chose these two
types of time reference because the failure modes of them are uncorrelated. A
datacenter typically consists of multiple \emph{time master} machines. Most of
these are GPS receivers with dedicated antennas. All remaining time master
machines use an atomic clock as reference. The references of all time masters are
compared against each other on a regular basis. To avoid incorrect time references
time masters also cross-check their local clock drift against the clock drift of
the reference. If these differ to much, the time master evicts itself.

Each machine has a \emph{timeslave daemon} installed which polls a variety of
time masters, mostly GPS masters but also some Armageddon masters. There is also
a mechanism to detect and reject liars which ensures that timeslave daemons only
synchronize to nonliars.

\begin{table}[ht]
  \centering
  \begin{tabular}{|l||l|}
    \hline
    {\bfseries Method} & {\bfseries Returns} \tabularnewline
    \hline\hline
    $TT.now()$ & $TTinterval$: [$earliest$, $latest$] \tabularnewline
    \hline
    $TT.after(t)$ & true if $t$ has definitely passed \tabularnewline
    \hline
    $TT.before(t)$ & true if $t$ has definitely now arrived \tabularnewline
    \hline
  \end{tabular}
  \caption{Methods provided by the TrueTime API}
  \label{tbl:truetime-api}
\end{table}

\section{Concurrency Control \tbr}
\label{sec:concurrency-control}

The TrueTime API allows Spanner to guarantee some correctness properties of the
implemented concurrency control. This enables the main features in terms of
transaction, i.e. externally-consistent transactions, lock-free read-only
transactions and so-called \emph{snapshot reads}. Snapshot reads are non-blocking
reads in the past at a specific timestamp $t$. Then, Spanner guarantees that
snapshot reads will see the effects of all transactions committed as of $t$.

Spanner provides different operations as can be seen in
Table~\ref{tbl:reads-writes}. Read-only transactions have to be made explicit.
This requires the application/user to predeclare that the read-only transaction
does not include write operations. Moreover, read-only transaction have the
benefits of snapshot isolation because they are \emph{lock-free}. Standalone
writes are simply read-write transaction without any read operations. Non-snapshot
standalone reads are implemented as read-only transactions. Read-only transactions
and snapshot reads proceed on any \emph{sufficiently up-to-date} replica. For
snapshot reads, the application/user has two options. On the one hand, a precise
timestamp can be specified at which the read should be executed. On the other
hand, an upper bound can be specified and Spanner choses a timestamp that
satisfies this bound.

\begin{table}
  \centering
  \begin{tabular}{|l||l|l|}
    \hline
    {\bfseries Operation} & {\bfseries Concurrency Control} &
    {\bfseries Replica required} \tabularnewline
    \hline\hline
    Read-Write transaction & Pessimistic & leader \tabularnewline
    \hline
    Read-Only transaction & Lock-free & leader; any
    \tabularnewline
    \hline
    Snapshot Read, client-provided timestamp & Lock-free & any \tabularnewline
    \hline
    Snapshot Read, client-provided bound & Lock-free & any \tabularnewline
    \hline
  \end{tabular}
  \caption{Reads and writes in Spanner}
  \label{tbl:reads-writes}
\end{table}

\subsection{Paxos Leader Leases \tbr}
\label{subsec:paxos-leader-leases}

The Paxos implementation of Spanner is designed for long-lived leaderships. As
usual in Paxos implementations, a leader is elected by a quorum of lease votes.
If a machine receives such a quorum of lease votes, it gets a lease (10 seconds
by default). The lease vote of a replica is extended automatically on successul
writes.

Spanner enforces a \emph{disjointness invariant} such that for each Paxos group,
each Paxos leader's \emph{lease interval} is disjoint from every other leader's.
In order to guarantee this invariant, Spanner constrains abdication of leaders.
It defines a maximum timestamp used by a leader, denoted as $s_{max}$, which is
advanced by several transactions. A leader must not abdicate before
$TT.after\left(s_{max}\right)$ has passed.

\subsection{Read-Write Transactions \tbr}
\label{subsec:read-write-transactions}

Read-write transaction use 2-phase locking and Spanner may only assign timestamps
in between the two phases. Spanner enforces another invariant, i.e. a
\emph{monotonicity invariant}, which makes use of the disjointness invariant.
This monotonicity invariant guarantees that within each Paxos group, Spanner
assigns timestamps to Paxos writes in monotonically increasing order. To guarantee
this across leaders, Spanner utilizes the disjointness invariant. This implies
that leaders must only assign timestamps within the interval of their respective
leader leases. $s_{max}$ is advanced to $s$ whenever a timestamp $s$ is assigned
by a leader.

To guarantee externally-consistent reads and writes, Spanner enforces a
\emph{external-consistency invariant}. In essence, this invariant guarantees that
whenever the commit of a transaction $T_1$, denoted as $e_1^{commit}$, occurs
before the start of a transaction $T_2$, denoted as $e_2^{start}$, then the commit
timestamp of $T_1$, denoted as $s_1$, is guaranteed to less than the timestamp
of $T_2$, denoted as $s_2$. More formally, the following must hold to guarantee
external consistency:
\begin{center}
$t_{abs}\left(e_1^{commit}\right) < t_{abs}\left(e_2^{start}\right) \Rightarrow s_1 < s_2$.
\end{center}

Spanner obeys two rules in order to guarantee this invariant:
\begin{description}
  \item[Start] Coordinator leader for a write $T_i$ assigns a commit timestamp
    $s_i \geq TT.now().latest$, where $TT.now()$ is executed right after the
    arrival event of commit request for $T_i$.
  \item[Commit Wait] Coordinator leader ensures that clients cannot see data
    committed by $T_i$ until $TT.after\left(s_i\right)$ is true. This ensures
    that the assigned timestamp $s_i$ is less than the absolute commit time of
    $T_i$.~\cite{Corbett:2012}
\end{description}

Figure~\ref{fig:read-write-transaction} illustrates the timeline of a distributed
read-write transaction. There are three participants groups involved, i.e. $P_2$,
$P_3$ and $P_{CL}$ (which is the coordinator leader). In the first step, all three
participants acquire their locks and execute $TT.now()$ right afterwards to obtain
timestamps independently. Every participant logs his timestamp through Paxos and
after this was done, sends its timestamp to the coordinator leader. The
coordinator leader then computes an overall timestamp for the transaction. This
timestamp must be greater than any timestamp the leader has assigned before,
greater or equal to all timestamps received by the participants and greater than
$TT.now().latest$ at the time the coordinator leader received the commit message.
Then, the computed timestamp is logged by the coordinator leader through Paxos
but no participants are notified until the commit wait rule is satisfied. After
commit wait is satisfied, the coordinator leader notifies all other participants
which can release their locks after receiving the notification.

\begin{figure}
  \centering
  \begin{tikzpicture}[thick, scale = 0.5]
    \node (left-end-t1) at (0,2) {};
    \node[right = 9 of left-end-t1] (right-end-t1) {};
    \draw[ultra thick, ->, >=stealth] (left-end-t1.center) -- (right-end-t1.center)
      node[right] (right-end-t1-t) {$t$};
    \node[left = 0 of left-end-t1.west] (left-end-t1-t) {$\mathbf{P_{CL}}$};

    \node[below = 1 of left-end-t1] (left-end-t2) {};
    \node[right = 9 of left-end-t2] (right-end-t2) {};
    \draw[->, >=stealth] (left-end-t2.center) -- (right-end-t2.center)
      node[right] (right-end-t2-t) {$t$};
    \node[left = 0 of left-end-t2.west] (left-end-t2-t) {$P_2$};

    \node[below = 1 of left-end-t2] (left-end-t3) {};
    \node[right = 9 of left-end-t3] (right-end-t3) {};
    \draw[->, >=stealth] (left-end-t3.center) -- (right-end-t3.center)
      node[right] (right-end-t3-t) {$t$};
    \node[left = 0 of left-end-t3.west] (left-end-t3-t) {$P_3$};

    \node[right = 0.25 of left-end-t1] (acquire-locks-t1) {};
    \node[right = 1.05 of left-end-t2] (acquire-locks-t2) {};
    \node[right = 0.65 of left-end-t3] (acquire-locks-t3) {};
    \foreach \x in {1, 2, 3}{
      \draw[blue] (acquire-locks-t\x.center) -- ++(0, -0.25);
      \draw[blue] (acquire-locks-t\x.center) -- ++(0, 0.25);
    }

    \node[above = 0.75 of acquire-locks-t1]
      (acquire-locks-t) {\scriptsize\color{blue}Acquire all locks};

    \foreach \x in {1, 2, 3}{
      \node[right = -0.1 of acquire-locks-t\x] (compute-s-t\x) {};
    }
    \node[below = 1 of acquire-locks-t3.center]
      (compute-s-t) {\scriptsize $s=TT.now()$};
    \draw[<-, >=stealth, dotted] (compute-s-t1.center) -- ++(0, -7);
    \draw[<-, >=stealth, dotted] (compute-s-t2.center) -- ++(0, -4.6);
    \draw[<-, >=stealth, dotted] (compute-s-t3.center) -- ++(0, -2);

    \node[right = 1 of compute-s-t2] (start-log-t2) {};
    \node[right = 1.75 of compute-s-t3] (start-log-t3) {};
    \node[above = 2 of start-log-t2] (start-log-t)
      {\scriptsize Start Logging};
    \draw[<-, >=stealth, dotted] (start-log-t2.center) -- ++(0, 4.25);
    \draw[<-, >=stealth, dotted] (start-log-t3.center) -- ++(0, 6.75);

    \node[right = 1 of start-log-t3] (end-log-t3) {};
    \node[right = 1.75 of start-log-t2] (end-log-t2) {};
    \node[above = 2 of end-log-t2] (end-log-t)
      {\scriptsize Logging done};
    \draw[<-, >=stealth, dotted] (end-log-t2.center) -- ++(0, 4.25);
    \draw[<-, >=stealth, dotted] (end-log-t3.center) -- ++(0, 6.75);
    \node[right = 5.5 of compute-s-t1] (compute-s-global) {};
    \draw[<-, >=stealth, black!50!green] (compute-s-global) ++(-0.1, 0) --
      (end-log-t3.north east);
    \draw[<-, >=stealth, black!50!green] (compute-s-global) ++(-0.65, 0) --
      (end-log-t2.north east);
    \draw[black!50!green, ultra thick] (compute-s-global.center) -- ++(0, -0.25);
    \draw[black!50!green, ultra thick] (compute-s-global.center) -- ++(0, 0.25);

    \node[above = 0.25 of compute-s-global] (compute-s-global-t)
      {\scriptsize\color{black!50!green} Overall $s$};
    \node[below = 0 of end-log-t3] (prepared-t)
      {\scriptsize\color{black!50!green}Prepared; Send $s$};

    \node[right = 0.5 of compute-s-global] (commit-wait-done) {};
    \node[below = 3.45 of commit-wait-done] (commit-wait-done-t)
      {\scriptsize Commit Wait done};
    \draw[<-, >=stealth, dotted] (commit-wait-done.center) -- ++(0, -7);
    \node[right = 0 of commit-wait-done] (release-locks-t1) {};
    \draw[red] (release-locks-t1.center) -- ++(0, -0.25);
    \draw[red] (release-locks-t1.center) -- ++(0, 0.25);
    \node[above right = 0.75 and -0.75 of release-locks-t1]
      (release-locks-t1-t) {\scriptsize\color{red}Release all locks};

    \node[right = -0.15 of release-locks-t1] (notify-t1) {};
    \node[right = 3.1 of end-log-t2] (notify-t2) {};
    \node[right = 3.35 of end-log-t3] (notify-t3) {};
    \foreach \x in {2, 3}{
      \draw[->, >=stealth, black!50!green] (notify-t1.center) --
        (notify-t\x.center); 
    }

    \node[right = -0.15 of notify-t2] (release-locks-t2) {};
    \node[right = -0.15 of notify-t3] (release-locks-t3) {};
    \draw[red] (release-locks-t2.center) -- ++(0, -0.25);
    \draw[red] (release-locks-t2.center) -- ++(0, 0.25);
    \draw[red] (release-locks-t3.center) -- ++(0, -0.25);
    \draw[red] (release-locks-t3.center) -- ++(0, 0.25);
  \end{tikzpicture}
  \caption{Read-write transaction}
  \label{fig:read-write-transaction}
\end{figure}

This also explains why, according to Table~\ref{tbl:reads-writes}, Read-write
transactions only involve the \emph{leader} replicas. This is because only the
leader replicas of each participant communicate with the coordinator leader.
Hence, no other replicas are involved.

\subsection{Snapshot Reads \tbr}
\label{subsec:reads-at-timestamp}

In order to provide snapshot reads, we first need to define
\emph{sufficiently up-to-date}. Therefore Spanner tracks a \emph{safe time},
denoted as $t_{safe}$, at each replica. $t_{safe}$ is defined as the maximum
timestamp at which the corresponding replica is up-to-date. Hence, a replica
satisfies reads at every timestamp $t$ if $t \leq t_{safe}$ holds. Spanner can
then perform a snapshot read at any replica that is sufficiently up-to-date. As
a consequence, snapshot reads have the same performance benefits as provided by
snapshot isolation since they are lock-free.

The safe time is determined by the safe time of a Paxos state machine, denoted
as $t_{safe}^{Paxos}$, and the safe t ime of the transaction manager, i.e.
$t_{safe}^{TM}$. The overall safe time of a replica then is the minimum of
these two. $t_{safe}^{Paxos}$ is simple equal to the time of the highest-applied
Paxos write because if the Paxos level is reached in the spanserver software stack,
then the corresponding transaction has already been committed. $t_{safe}^{TM}$
is $\infty$ if there exist distributed transaction which are in between the two
phases of two phases of the 2-phase commit because in this case the system does
not guaranteed that this transaction will commit.

\subsection{Read-Only Transactions \tbr}
\label{subsec:read-only-transactions}

Read-only transactions are also lock-free and execute in two phases. In the first
phase, a timestamp $s_{read}$ is assigned. In the second phase, the read-only
transaction is executed as a snapshot read at $s_{read}$. Spanner assigns
the oldest timestamp which preserves external consistency to $s_{read}$.

In general, read-only transactions involve group leaders and they may also involve
any other replica which is sufficiently up-to-date. Group leaders are involved
in the negotiation phase. In this phase, all involved Paxos groups negotiate in
order to determine $s_{read}$. To accomplish this, Spanner requires every
read-only transaction to define a \emph{scope} expression which summarizes all
keys that will be read by the respective transaction.

If all values of the scope can be served by a single Paxos group, the client
issues the read-only transaction to the corresponding group leader. The group
leader then assign the timestamp of the last committed write to $s_{read}$ and
executes the transaction.

If serving the values of the scope involves multiple Paxos groups, Spanner avoids
a negotiation round (which would be one viable options). Instead, Spanner simply
assigns$TT.now().latest$ to $s_{read}$, even though this may have to wait for
$t_{safe}$ to advance.

In both cases, the snapshot read executed at $s_{read}$ can execute on any
replica. The leader replicas are only involved in the negotiation phase.

\subsection{Schema-Change Transactions \tbr}
\label{subsec:schema-change-transactions}

Schema-change transactions are variants of standard transactions which change
schemas atomically without blocking the whole system.

There are two phases involved. In the first phase, the \emph{prepare} phase,
Spanner explicitly assigns a future timestamp to the schema-change transaction.
In the phase, reads and writes synchronize with any registered schema-change
timestamp at time $t$. Of course, this concerns only the reads and writes which
implicitly depend on this schema. Any read/write transaction may proceed if
its timestamp preceeds $t$. If its timestamp is after $t$, the corresponding
read/write has to block behind the schema-change transaction. This is essentially
enabled by the TrueTime API because without TrueTime, it would not be guaranteed
that the timestamp of the schema-change transaction is in the future.

\section{Performance \tbr}
\label{sec:performance}

\begin{table}[ht]
  \centering
  { \footnotesize
    \begin{tabular}{|l||l|l|l||l|l|l|}
      \hline
      & \multicolumn{3}{c||}{{\bfseries Latency in [ms] (less is better)}} &
      \multicolumn{3}{c|}{{\bfseries Throughput in [Kops/sec] (more is better)}}
      \tabularnewline
      \cline{2-7}
      {\bfseries Replicas} & {\bfseries Write} & {\bfseries RO Transaction} &
      {\bfseries Snapshot Read} & {\bfseries Write} &
      {\bfseries RO Transaction} & {\bfseries Snapshot Read} \tabularnewline
      \hline\hline
      1D & $9.4\pm0.6$ & - & - & $4.0\pm0.3$ & - & - \tabularnewline
      \hline
      1 & $14.4\pm1.0$ & $1.4\pm0.1$ & $1.3\pm0.1$ & $4.1\pm0.05$ &
      $10.9\pm0.4$ & $13.5\pm0.1$ \tabularnewline
      \hline
      3 & $13.9\pm0.6$ & $1.3\pm0.1$ & $1.2\pm0.1$ & $2.2\pm0.5$ &
      $13.8\pm3.2$ & $38.5\pm0.3$ \tabularnewline
      \hline
      5 & $14.4\pm0.4$ & $1.4\pm0.05$ & $1.3\pm0.04$ & $2.8\pm0.3$ &
      $25.3\pm5.2$ & $50.0\pm1.1$ \tabularnewline
      \hline
    \end{tabular}
  }
  \caption{Performance of operations. Mean and standard deviation over 10
    runs.~\cite{Corbett:2012}
  }
  \label{tbl:performance-of-operations}
\end{table}

This section summarizes the core results of the evaluation of Spanner, i.e. the
latency and throughput performance as well as some results about availability
and the TrueTime API.

Table~\ref{tbl:performance-of-operations} summarizes the latency and throughput
performance of Spanner. 1D means one replica with commit wait disabled. Hence,
only write transactions are evaluated because commit wait is only involved in such
transactions. As we can see, the latency of all transaction types is almost stable
for an increasing number of replicas whereas the standard deviation decreases.
The throughput of read-only transaction increases with the number of replicas
because the number of spanservers increases (in the experiments, the number of
replicas and the number of spanservers were the same. The throughput of write
transactions also benefits from this fact. However, the number of write operations
to perform increases linearly with the number of replicas. Snapshot reads benefit
the most from the Spanner architecture. The throuhgput of snapshot reads increases
almost linear with the number of replicas.

The availability of Spanner is only affected if zones containing multiple leaders
are shut down without notifying the leader upfront. In this case, it takes Spanner
about 10s to recover full availability. If the leaders are notified to abdicate
before their zone is shut down, the availability stays almost unaffected. The same
if true if a zone without a leader is shut down.

Google also provided results for some experiments concerning TrueTime. All
assumptions Google made when designing Spanner would break if the local clock
drift would exceed $200\mu$s/sec. But Google found that clock issues are extremely
infrequent relative to more serious hardware problems, e.g. bad CPUs. Moreover,
Google provided empirical results on the values of $\epsilon$. Typically,
$\epsilon$ is never larger than 10ms. It decreases as the network infrastructure
is improved and increases if time masters go offline.

\section{Future Work \tbr}
\label{sec:future-work}

Google plans to implement optimistic parallel reads but it turns out that
implementing this is non-trivial. They also want to support direct configuration
changes of the Paxos state machines and to reduce $\epsilon$ to be $\leq 1ms$ to
improve the performance if data is sharded between nearby datacenters.

Moreover, the node-local performance can be improved for complex queries. By now,
node-local data structures are designed for key-value accesses and these could be
improved by algorithms and data structure from database literature.

To improve the performance even further, Google plans to move processes between
datacenters if the data they operate on is moved. This would have a high impact
because the performance of the processes is close to optimal if the data is as
local as possible. However, moving processes between datacenters raises even more
complex problems.

\section{Discussion \tbr}
\label{sec:discussion}

The system introduced by Google is definitly impressive because they were able
to establish a notion of global wall clock time (that's how they call it in their
presentation) which enabled them to guarantee serialization order for
globally-distributed transactions. To the best of the author's knowledge, Spanner
was the first system to provide much stronger consistency and time semantics while
data is sharded and replicated at global scale. Google also proved all
'\emph{Relational databases are dead}' people wrong even though systems like
Bigtable helped creating the NoSQL trend. Reading about a system like Spanner is
definitly more than a vital sign of relational database systems. From a
developer's point of view, this is definitly interesting because Spanner provides
the more intuitive schematized view of tables. In contrast to NOSQL database
systems, Spanner could be used as replacement for other relational database
systems, e.g. MySQL, directly. Spanner is also highly flexible to schema changes
because of the atomic schema change transactions.

However, I believe that rebuilding a Spanner-like system is difficult for
small- to mid-sized companies because there is a lot of hardware involved, for
example all the GPS and atomic clocks. Unless this hardware infrastructure is
not available it is also hard to reimplement the TrueTime API. Nonetheless,
companies which need a globally-distributed database with all the guarantees
Spanner provides are likely to be almost as big as Google. Hence, such companies
may also be able to build a Spanner-like system. If every datacenter would provide
GPS and Armageddon time masters independently, the reimplementation of such a
system would be much easier.

Another part that may influence the performance of Spanner is the distributed
file system, i.e. Colossus, the successor of the Google File System (GFS). I did
not find reliable information about the implementation or the performance of
Colossus. It is unknown how much impact this distributed file system has on the
performance of Spanner. Thus, it is hard to say if a reimplementation of Spanner
would be able to compete with Spanner's performance.

Of course, such a system may also be built upon e.g. the Network Time Protocol
(NTP). But this would result in poor performance compared to Spanner because the
error interval would be much larger compared to $2\cdot\epsilon$. And besides
providing distributed general-purpose transactions, the performance of Spanner
makes the whole system that impressive. A system using e.g. NTP would never be
able to achieve this performance while guaranteeing serialization order. Spanner
can synchronize multiple systems without any communication between these systems.
NTP-based systems would have to communicate with each other. For systems far apart
such a system would never be able to compete with Spanner.

Spanner is robust to GPS blackouts because of the usage ofatomic clocks, even
though the performance of Spanner may suffer due to high values of $\epsilon$.
Would be interesting to see how such a blackout affects Spanner's performance.

It would also be interesting how failures of the two singletons per universe,
i.e. universemaster and placement driver, affect Spanner. Since the universemaster
is just a debugging console, most probably a failure in this component will not
affect Spanner at all. This would probably be a totally different story if the
placement driver does not operate correctly. First of all, there would be no
load balancing between zones. Hence, zones could be run out of space if the
placement driver is down long enough. Furthermore, updated replication
constraints, would not be propagated since the placement driver is responsible
to move the data if the replication constraints are modified.

In terms of applications, applications which read a lot but write rarely or little
data would benefit the most from a system like Spanner. If an application uses
lots of replicas which are spread across the globe and writes large amounts of data
regularly, Spanner may not provide the performance mentioned in the original
paper. However, Spanner may still outperform both, Megastore and Bigtable while
providing much stronger consistency guarantees.

%\clearpage
\bibliographystyle{abbrv}
\bibliography{paper}

\end{document}